# Communicable, Maternal, Neonatal, and Nutritional Disease Burden (CMNN)

```{r,echo = FALSE,include = FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the communicable diseases data
data_cmnn <- read_csv(here("data", "burden-of-disease-cmnn.csv"))
```

## Table of Estimates for CMNN Burden Over Time

```{r, echo = FALSE}
# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.

# Filter for only Switzerland, Canada, and South Africa
cmnn_countries_dalys <- data_cmnn %>%
  filter(Entity %in% c("Switzerland", "Canada", "South Africa"))

# Pivot the data so that each country's DALYs are shown in separate columns
cmnn_pivoted_data <- cmnn_countries_dalys %>%
  pivot_wider(names_from = Entity,Year, values_from = `DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`)

# Generate a table using kable from the knitr package
kable(cmnn_pivoted_data, format = "html", caption = "DALY Burden Over Time for Switzerland, Canada, and South Africa")
```

## Plot Showing Trends in CMNN Burden Over Time

```{r, echo = FALSE}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
# Convert the wide format back to long format for ggplot
cmnn_long_data <-cmnn_pivoted_data %>%
  pivot_longer(cols = c("Switzerland", "Canada", "South Africa"),
               names_to = "Entity",
               values_to = "DALYs")
# Create a line plot with ggplot2
ggplot(cmnn_long_data, aes(x = Year, y = DALYs, color = Entity)) +
  geom_line() + 
  scale_x_continuous(breaks = seq(min(cmnn_long_data$Year), max(cmnn_long_data$Year), by = 5)) +
  labs(title = "Trend of DALY Burden Over Time",
       x = "Year",
       y = "DALYs",
       color = "Entity") +
  theme_minimal() +
  scale_color_manual(values = c("Switzerland" = "blue", "Canada" = "red", "South Africa" = "green"))

```

## Summary of CMNN Burden Findings

There is a noticeable difference in the trend lines between the countries. For instance, South Africa shows a higher and more volatile DALY burden over the years compared to Switzerland and Canada, indicating a greater impact of communicable, maternal, neonatal, and nutritional diseases.In 2005 there's a notable spikes in South Africa highlight periods of health crises or gaps in disease control and prevention, in this case HIV/AIDs).
