# Non-Communicable Disease Burden (NCD)

```{r, echo = FALSE,include = FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here, kable)


# Import the non-communicable diseases data
data_ncd <- read_csv(here("data", "burden-of-disease-ncd.csv"))
view(data_ncd)
```

## Table of Estimates for NCD Burden Over Time

```{r, echo = FALSE}
# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
tableCAN_SA_CH<-data_ncd %>% 
  filter(Entity%in%c("Canada", "South Africa", "Switzerland")) %>% 
  pivot_wider(names_from="Entity", "Year",
              values_from="DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)")

  # Use kable() from the knitr package to render the table.
kable(tableCAN_SA_CH, format="html", caption="DALYs per Year for Canada, South Africa and Switzerland")
```

## Plot Showing Trends in NCD Burden Over Time 

```{r, echo = FALSE}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
dataCAN_SA_CH<-tableCAN_SA_CH %>% 
  pivot_longer(cols=2:4, names_to="Country", values_to="DALY")

#plotCAN_SA_CA <- 
  ggplot(dataCAN_SA_CH,
        aes(x=Year, y=DALY, color=Country))+
  geom_line(size=1)+
    labs(title = "Trend of DALYs per year attributable to non-communicable diseases",
         subtitle ="Time range 1990-2019",
         y="DALYs (Age-standardized)")+
    scale_x_continuous(breaks=seq(1990,2020,2))

```

## Summary of NCD Burden Findings

In the data we see an overall decline in disability-adjusted life years in all three countries. While this decline has been continously observable since 1990 for Switzerland, Canada has stable numbers since 2015 after an initial decline. In contrast, South Africa has seen an initial increase between 1990 and 2000 with a continous decline since then.


